Realizability-Preserving DG-IMEX Method for the Two-Moment Model of Fermion Transport \tnoteref{support}

### 2018-06-20 Documents/M1/realizableFermionicM1/sections/appendix.tex
Modified the appendix: modified the consistency requirement and 2nd-order streaming requirement.

It can be shown that no 2-stage positivity-preserving IMEX exist \cite{hu_etal_2018}.
Hence we considered 3-stage IMEX scheme.
There are two kinds of implicit DIRK scheme, Type A and Type ARS.
Here we discuss the simpler case, Type ARS, first.
The Butcher tableau for ARS type IMEX scheme is
\begin{equation}
  \begin{array}{c | c c c}
  	         0           & 0                    & 0                    & 0                    \\
  	\tilde{c}_{2} & \tilde{a}_{21}       & 0                    & 0                    \\
  	\tilde{c}_{3} & \tilde{a}_{31}       & \tilde{a}_{32}       & 0                    \\ \hline
  	                     & \tilde{w}_{1} & \tilde{w}_{2} & \tilde{w}_{3}
  \end{array}
  \qquad
  \begin{array}{c | c c c}
  	     0       & 0            & 0            & 0            \\
  	c_{2} & 0            & a_{22}       & 0            \\
  	c_{3} & 0            & a_{32}       & a_{33}       \\ \hline
  	             & w_{1} & w_{2} & w_{3}
  \end{array}.
\end{equation}
The properties a desired IMEX scheme should have are: 2nd-order accurate in the streaming, globally stiffly accurate, positivity-preserving, diffusion accurate and an affordable time-step restriction.
Then the problem that find such a scheme can be rewritten as an optimization problem:\\
Find $(\tilde{a}_{21}, \tilde{a}_{31}, \tilde{a}_{32}, \tilde{w}_{1}, \tilde{w}_{2}, \tilde{w}_{3}, a_{22}, a_{32}, a_{33}, w_{1}, w_{2}, w_{3})$ such that
\begin{align*}
\text{maximize} ~ c_{sch} = &\min \{ \dfrac{c_{20}}{\tilde{c}_{20}}, \dfrac{c_{30}}{\tilde{c}_{30}}, \dfrac{c_{32}}{\tilde{c}_{32}}\}\\
 \text{under these constraints:} \qquad & \\
 \text{(GSA)} \qquad\ &\tilde{w}_{i} = \tilde{a}_{3i}, \quad w_{i} = a_{3i}, ~\text{for} ~ i = 1,2,3\\
 \text{(2nd-order accuracy)} \qquad\ &\sum_{i=1}^{3}  \tilde{w}_{i} = 1, \quad \sum_{i=1}^{3}  \tilde{w}_{i}\tilde{c}_{i} = 1/2, \\
 \text{(positivity-preserving)} \qquad & a_{ii} > 0, c_{i0} \geq 0, \tilde{c}_{i0} \geq 0, ~\text{for} ~ i = 2,3 \\
 & c_{32} \geq 0, \tilde{c}_{32} \geq 0, \\
 \text{(diffusion accurate)} \qquad & e^{T}_iA^{-1}\tilde{A}e = 1, ~\text{for} ~ i = 2,3, \\
 \text{(affordable)} \qquad & c_{sch} > 0.
\end{align*}
Imposing those equality constraints (including all the constraints exclude those from the positivity-preserving and affordable requirements), the Butcher tableau can be simplified as:
\begin{equation}
  \begin{array}{c | c c c}
  	     0       & 0            & 0 & 0 \\
  	\frac{1}{2x} & \frac{1}{2x} & 0 & 0 \\
  	     1       & 1-x          & x & 0 \\ \hline
  	             & 1-x          & x & 0
  \end{array}
  \qquad
  \begin{array}{c | c c c}
  	     0       & 0 & 0            & 0 \\
  	\frac{1}{2x} & 0 & \frac{1}{2x} & 0 \\
  	     1       & 0 & 1-y          & y \\ \hline
  	             & 0 & 1-y          & y
  \end{array}.
\end{equation}
And the rest inequalities and the problem were left to be: \\
Find $(x,y)$ such that
\begin{align*}
\text{maximize} ~ c_{sch} = &\min \{ 2x, \dfrac{1-2x(1-y)}{y-x}, 2(1-y)\}\\
 \text{under these constraints:} \qquad & x > 0, ~ 1 \geq y > 0, \\
 & x \leq y, ~\dfrac{1}{2x} \geq 1-y.
\end{align*}
A solution family is $x=\frac{1}{2}$, $y=\frac{1}{2} + \epsilon$ with $\epsilon \in [0, \frac{1}{2})$ and $c_{sch} = 1 - 2\epsilon$.
The PARSD scheme follows.
It's clear that $\epsilon = 0$ and $c_{sch} = 1$ is the most optimized solution.

@@

Proof of GSA IMEX schemes of Type~A with 3 stages and the following properties does not exist
\begin{itemize}
  \item Second-order in the streaming limit;
  \item Well-behaved in the diffusion limit;
  \item Positivity-preserving.
\end{itemize}
First of all, the GSA IMEX scheme of Type A with 3 stages have the following form:
\begin{equation*}
  \begin{array}{c | c c c}
  	      0       & 0              & 0              & 0 \\
  	\tilde{c}_{2} & \tilde{a}_{21} & 0              & 0 \\
  	\tilde{c}_{3} & \tilde{a}_{31} & \tilde{a}_{32} & 0 \\ \hline
  	              & \tilde{a}_{31} & \tilde{a}_{32} & 0
  \end{array}
  \qquad
  \begin{array}{c | c c c}
    c_{1} & a_{11} & 0      & 0      \\
  	c_{2} & a_{21} & a_{22} & 0      \\
  	c_{3} & a_{31} & a_{32} & a_{33} \\ \hline
  	      & a_{31} & a_{32} & a_{33}
  \end{array}.
\end{equation*}
The second-order in the streaming limit requires
\begin{align}
\tilde{a}_{31} + \tilde{a}_{32} = 1, \quad \tilde{a}_{32}\cdot \tilde{a}_{21} =  \frac{1}{2}.
\end{align}
An implicit property, the first-order in general transport condition, requires
\begin{align}
 \quad a_{31} + a_{32} + a_{33} = 1.
 \label{eq:1stOrderGeneral}
\end{align}
The diffusion accurate Eq.~\eqref{eq:diffusionCondition} requires 
\begin{align}
\tilde{a}_{21} = a_{22}, \quad -\dfrac{a_{32} \tilde{a}_{21}}{a_{22}a_{33}} + \dfrac{1}{a_{33}} =1.
\label{eq:diffusionAccurate}
\end{align}
Substituting Eq.~\eqref{eq:1stOrderGeneral} into Eq.~\eqref{eq:diffusionAccurate} gives
\begin{align}
\tilde{a}_{21} = a_{22}, \quad a_{31} = 0.
\label{eq:a31is0}
\end{align}
Positivity-preserving requires
\begin{align*}
& a_{11} > 0, ~ a_{22} >0, ~ a_{33}>0,\\
& c_{20} = 1 - \frac{a_{21}}{a_{11}} \geq 0, ~ c_{30} = 1-\frac{a_{31}}{a_{11}} + \frac{a_{32}a_{21}}{a_{22}a_{11}} - \frac{a_{32}}{a_{22}}\geq 0, \\
& c_{21} = \frac{a_{21}}{a_{11}} \geq 0, ~ c_{31} = \frac{a_{31}}{a_{11}} - \frac{a_{32}a_{21}}{a_{22}a_{11}} \geq 0, ~ c_{32} = \frac{a_{32}}{a_{22}} \geq 0,\\
& \tilde{c}_{21} = \tilde{a}_{21} \geq 0, \tilde{c}_{31} = \tilde{a}_{31} - \frac{a_{32}\tilde{a}_{21}}{a_{22}} \geq 0, \tilde{c}_{32} = \tilde{a}_{32} \geq 0.
\end{align*}
Note that since $a_{31} = 0$, $a_{21} \geq 0$, $a_{32} \geq 0$, $a_{22} > 0$ and $a_{11} >0$, $c_{31} \geq 0$ only holds for $c_{31} = 0$.
Unfortunately, $c_{31} = 0$ will drag $c_{\mbox{\tiny Sch}} = 0$ since $c_{\mbox{\tiny Sch}} = \min \left\lbrace \dfrac{c_{21}}{\tilde{c}_{21}}, \dfrac{c_{31}}{\tilde{c}_{31}}, \dfrac{c_{32}}{\tilde{c}_{32}}\right\rbrace $.
Therefore, GSA IMEX schemes of Type~A with 3 stages and the above properties does not exist.

>>>>

It can be shown that no 2-stage positivity-preserving IMEX exist \cite{hu_etal_2018}.
Hence we considered 3-stage IMEX scheme.
There are two kinds of implicit DIRK scheme, Type A and Type ARS.
Here we discuss the simpler case, Type ARS, first.
The Butcher tableau for ARS type IMEX scheme is
\begin{equation}
  \begin{array}{c | c c c}
  	         0           & 0                    & 0                    & 0                    \\
  	\tilde{c}_{2} & \tilde{a}_{21}       & 0                    & 0                    \\
  	\tilde{c}_{3} & \tilde{a}_{31}       & \tilde{a}_{32}       & 0                    \\ \hline
  	                     & \tilde{w}_{1} & \tilde{w}_{2} & \tilde{w}_{3}
  \end{array}
  \qquad
  \begin{array}{c | c c c}
  	     0       & 0            & 0            & 0            \\
  	c_{2} & 0            & a_{22}       & 0            \\
  	c_{3} & 0            & a_{32}       & a_{33}       \\ \hline
  	             & w_{1} & w_{2} & w_{3}
  \end{array}.
\end{equation}
The properties a desired IMEX scheme should have are: 2nd-order accurate in the streaming, globally stiffly accurate, positivity-preserving, diffusion accurate and an affordable time-step restriction.
Then the problem that find such a scheme can be rewritten as an optimization problem:\\
Find $(\tilde{a}_{21}, \tilde{a}_{31}, \tilde{a}_{32}, \tilde{w}_{1}, \tilde{w}_{2}, \tilde{w}_{3}, a_{22}, a_{32}, a_{33}, w_{1}, w_{2}, w_{3})$ such that
\begin{align*}
\text{maximize} ~ c_{sch} = &\min \{ \dfrac{c_{20}}{\tilde{c}_{20}}, \dfrac{c_{30}}{\tilde{c}_{30}}, \dfrac{c_{32}}{\tilde{c}_{32}}\}\\
 \text{under these constraints:} \qquad & \\
 \text{(GSA)} \qquad\ &\tilde{w}_{i} = \tilde{a}_{3i}, \quad w_{i} = a_{3i}, ~\text{for} ~ i = 1,2,3\\
 \text{(consistency)} \qquad\ & \sum_{i=1}^{3}  \tilde{w}_{i} = 1, \quad \sum_{i=1}^{3} w_{i} = 1\\ 
 \text{(2nd-order streaming accuracy)} \qquad\ & \sum_{i=1}^{3}  \tilde{w}_{i}\tilde{c}_{i} = 1/2, \\
 \text{(positivity-preserving)} \qquad & a_{ii} > 0, c_{i0} \geq 0, \tilde{c}_{i0} \geq 0, ~\text{for} ~ i = 2,3 \\
 & c_{32} \geq 0, \tilde{c}_{32} \geq 0, \\
 \text{(diffusion accurate)} \qquad & e^{T}_iA^{-1}\tilde{A}e = 1, ~\text{for} ~ i = 2,3, \\
 \text{(affordable)} \qquad & c_{sch} > 0.
\end{align*}
Imposing those equality constraints (including all the constraints exclude those from the positivity-preserving and affordable requirements), the Butcher tableau can be simplified as:
\begin{equation}
  \begin{array}{c | c c c}
  	     0       & 0            & 0 & 0 \\
  	\frac{1}{2x} & \frac{1}{2x} & 0 & 0 \\
  	     1       & 1-x          & x & 0 \\ \hline
  	             & 1-x          & x & 0
  \end{array}
  \qquad
  \begin{array}{c | c c c}
  	     0       & 0 & 0            & 0 \\
  	\frac{1}{2x} & 0 & \frac{1}{2x} & 0 \\
  	     1       & 0 & 1-y          & y \\ \hline
  	             & 0 & 1-y          & y
  \end{array}.
\end{equation}
And the rest inequalities and the problem were left to be: \\
Find $(x,y)$ such that
\begin{align*}
\text{maximize} ~ c_{sch} = &\min \{ 2x, \dfrac{1-2x(1-y)}{y-x}, 2(1-y)\}\\
 \text{under these constraints:} \qquad & x > 0, ~ 1 \geq y > 0, \\
 & x \leq y, ~\dfrac{1}{2x} \geq 1-y.
\end{align*}
A solution family is $x=\frac{1}{2}$, $y=\frac{1}{2} + \epsilon$ with $\epsilon \in [0, \frac{1}{2})$ and $c_{sch} = 1 - 2\epsilon$.
The PARSD scheme follows.
It's clear that $\epsilon = 0$ and $c_{sch} = 1$ is the most optimized solution.

@@

Proof of GSA IMEX schemes of Type~A with 3 stages and the following properties does not exist
\begin{itemize}
  \item Second-order in the streaming limit;
  \item Well-behaved in the diffusion limit;
  \item Positivity-preserving.
\end{itemize}
First of all, the GSA IMEX scheme of Type A with 3 stages have the following form:
\begin{equation*}
  \begin{array}{c | c c c}
  	      0       & 0              & 0              & 0 \\
  	\tilde{c}_{2} & \tilde{a}_{21} & 0              & 0 \\
  	\tilde{c}_{3} & \tilde{a}_{31} & \tilde{a}_{32} & 0 \\ \hline
  	              & \tilde{a}_{31} & \tilde{a}_{32} & 0
  \end{array}
  \qquad
  \begin{array}{c | c c c}
    c_{1} & a_{11} & 0      & 0      \\
  	c_{2} & a_{21} & a_{22} & 0      \\
  	c_{3} & a_{31} & a_{32} & a_{33} \\ \hline
  	      & a_{31} & a_{32} & a_{33}
  \end{array}.
\end{equation*}
The consistency of the discrete system to the continuous system requires
\begin{align}
 \tilde{a}_{31} + \tilde{a}_{32} = 1, \quad \quad a_{31} + a_{32} + a_{33} = 1.
 \label{eq:1stOrderGeneral}
\end{align}
The second-order in the streaming limit requires
\begin{align}
\tilde{a}_{32}\cdot \tilde{a}_{21} =  \frac{1}{2}.
\end{align}
The diffusion accurate Eq.~\eqref{eq:diffusionCondition} requires 
\begin{align}
\tilde{a}_{21} = a_{22}, \quad -\dfrac{a_{32} \tilde{a}_{21}}{a_{22}a_{33}} + \dfrac{1}{a_{33}} =1.
\label{eq:diffusionAccurate}
\end{align}
Substituting the second equation in Eq.~\eqref{eq:1stOrderGeneral} into Eq.~\eqref{eq:diffusionAccurate} gives
\begin{align}
\tilde{a}_{21} = a_{22}, \quad a_{31} = 0.
\label{eq:a31is0}
\end{align}
Positivity-preserving requires
\begin{align*}
& a_{11} > 0, ~ a_{22} >0, ~ a_{33}>0,\\
& c_{20} = 1 - \frac{a_{21}}{a_{11}} \geq 0, ~ c_{30} = 1-\frac{a_{31}}{a_{11}} + \frac{a_{32}a_{21}}{a_{22}a_{11}} - \frac{a_{32}}{a_{22}}\geq 0, \\
& c_{21} = \frac{a_{21}}{a_{11}} \geq 0, ~ c_{31} = \frac{a_{31}}{a_{11}} - \frac{a_{32}a_{21}}{a_{22}a_{11}} \geq 0, ~ c_{32} = \frac{a_{32}}{a_{22}} \geq 0,\\
& \tilde{c}_{21} = \tilde{a}_{21} \geq 0, \tilde{c}_{31} = \tilde{a}_{31} - \frac{a_{32}\tilde{a}_{21}}{a_{22}} \geq 0, \tilde{c}_{32} = \tilde{a}_{32} \geq 0.
\end{align*}
Note that since $a_{31} = 0$, $a_{21} \geq 0$, $a_{32} \geq 0$, $a_{22} > 0$ and $a_{11} >0$, $c_{31} \geq 0$ only holds for $c_{31} = 0$.
Unfortunately, $c_{31} = 0$ will drag $c_{\mbox{\tiny Sch}} = 0$ since $c_{\mbox{\tiny Sch}} = \min \left\lbrace \dfrac{c_{21}}{\tilde{c}_{21}}, \dfrac{c_{31}}{\tilde{c}_{31}}, \dfrac{c_{32}}{\tilde{c}_{32}}\right\rbrace $.
Therefore, GSA IMEX schemes of Type~A with 3 stages and the above properties does not exist.
